Predicting tree failure likelihood for utility risk mitigation via artificial intelligence

### 2021-03-05 docs/manuscript/manuscript.tex
final edits to first journal submission

\eject
\section{Conclusion}
We have demonstrated the efficacy of an artificial intelligence framework for predicting tree failure likelihood
(\textit{Probable}, \textit{Possible} and \textit{Improbable}) with respect to utility infrastructure. Specifically, we
developed a convolutional neural network with state-of-the-art configurations.  We applied data augmentation and
pre-processing strategies to increase the size of our dataset by a factor of 5, thus generating 2525 images.  From an
initial resolution of $4032 \times 3024$ pixels, we created three sets of image resolutions: $64 \times 64$,
$128 \times 128$ and $224 \times 224$. We then defined four classification scenarios to investigate the performance of
various groupings of the three categories: Pr\_Im (\textit{Probable} vs.\ \textit{Improbable}); PrPo\_Im
(\textit{Probable} and \textit{Possible} vs.\ \textit{Improbable}); Pr\_PoIm (\textit{Probable} vs.\ \textit{Possible}
and \textit{Improbable}); and Pr\_Po\_Im (\textit{Probable} vs.\ \textit{Possible} vs. \textit{Improbable}).
We then optimized eight of our CNN hyperparameters for 12 classification
scenario and image resolution combinations.  We conducted five-fold cross-validation for each of the 12 cases, and
assessed model performance based on accuracy and the macro-averages of precision, recall and $F_{1}$ score.

@@

\section{Data Availability Statement}
All data, models, or code generated or used during the study are available in a GitHub repository online at \url{https://github.com/narslab/tree-risk-ai}.

@@

\appendix
\section{Summary of validation performance metrics}
For further reference we tabulate the mean and standard deviation of the performance metrics obtained on the scenario-resolution sensitivity tests. These values are shown in \autoref{tab:meanstd}.
The metrics are accuracy, macro-average $F_1$ ($F_1^m$), macro-average recall ($Re^m$) and macro-average precision ($Pr^m$).

@@

\section{Welch $t$-test results}
Here, we summarize the p-values resulting from the Welch independent $t$-tests used to determine if the mean values of the various performance metrics are statistically significant or not.
Strong evidence of significant difference is given by a low p-value (a typical threshold is $< 0.01$).

>>>>

\section{Conclusion}
We have demonstrated the efficacy of an artificial intelligence framework for predicting tree failure likelihood
(\textit{Probable}, \textit{Possible} and \textit{Improbable}) with respect to utility infrastructure. Specifically, we
developed a convolutional neural network with state-of-the-art configurations.  We applied data augmentation and
pre-processing strategies to increase the size of our dataset by a factor of 5, thus generating 2525 images.  From an
initial resolution of $4032 \times 3024$ pixels, we created three sets of image resolutions: $64 \times 64$,
$128 \times 128$ and $224 \times 224$. We then defined four classification scenarios to investigate the performance of
various groupings of the three categories: Pr\_Im (\textit{Probable} vs.\ \textit{Improbable}); PrPo\_Im
(\textit{Probable} and \textit{Possible} vs.\ \textit{Improbable}); Pr\_PoIm (\textit{Probable} vs.\ \textit{Possible}
and \textit{Improbable}); and Pr\_Po\_Im (\textit{Probable} vs.\ \textit{Possible} vs. \textit{Improbable}).
We then optimized eight of our CNN hyperparameters for 12 classification
scenario and image resolution combinations.  We conducted five-fold cross-validation for each of the 12 cases, and
assessed model performance based on accuracy and the macro-averages of precision, recall and $F_{1}$ score.

@@

\section{Data Availability Statement}
All data, models, or code generated or used during the study are publicly available in a GitHub repository online at \url{https://github.com/narslab/tree-risk-ai}.

@@

\appendix
\section{Summary of validation performance metrics}
For further reference we tabulate the mean and standard deviation of the performance metrics obtained on the scenario-resolution sensitivity tests. These values are shown in \autoref{tab:meanstd}.
The metrics are: accuracy, macro-average $F_1$ ($F_1^m$), macro-average recall ($Re^m$) and macro-average precision ($Pr^m$).

@@

\section{Welch $t$-test results}
Here, we summarize the p-values resulting from the Welch independent $t$-tests used to determine if the mean values of the various performance metrics are statistically significant or not.
Strong evidence of significant difference is given by a low p-value (a typical threshold is $< 0.01$).
In \autoref{tab:welch-resolution}, we consider pairwise comparisons between different image resolutions in all four scenarios.
In \autoref{tab:welch-scenario}, we compare different scenarios for the same image resolution.
